### 2013-07-31 Tex/chapters/results.tex
Vorzeichen der Kritischen Exponenten korrigiert (hoffentlich)
finite size scaling und critical exponents  ausführlicher

\subsection{Finite Size Effects}
\label{ssec:finitesize}
    The aim of this thesis is to find the critical temperature \(T_c\)
    of the disordered Ising model in dependence of the disorder parameter
    \(\sigma\). At \(T_c\) the mean absolute magnetization \(\avg{|m|}\) of
    the system will show a steep decline to zero and the susceptibility
    \(\chi = \frac{N}{T}\brac{\avg{m^2}-\avg{m}^2}\)
    will diverge. In fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM_L}
    it is easy to see that the steep decline of \(\avg{|m|}\)
    occurs at lower temperatures \(T\) for higher
    disorder parameters \(\sigma\) (here with a RNG).
    \begin{figure}[htbp]
        \centering
        \subfigure[Dependence of the Phase Transition on $\sigma$][]
        {
            \label{sfig:smeared_out:meanM_L}
            \includegraphics[width=0.45\textwidth]{plots/Mean_M_L_128}
        }
        \subfigure[Example for Finite Size Effects][]
        {
            \label{sfig:smeared_out:meanM}
            \includegraphics[width=0.45\textwidth]{plots/meanM}
        }
        \caption[Phase Transition and Finite Size Effects]
        {
            \subref{sfig:smeared_out:meanM_L}: Effects of the disorder
            parameter $\sigma$ on the phase transition
            for an underlying RNG with $L=128$.\\
            \subref{sfig:smeared_out:meanM}: Effects of different system
            sizes at \(\sigma = 0\). Dotted lines are guides to the eye.
        }
        \label{fig:smeared_out}
    \end{figure}\\
    But in the shown plots there occurs no steep decline to zero, but a
    smooth decline. \(\avg{\abs{m}}(T_c) = 0\) is only present in infinite
    systems, hence no computer simulation will show the exact behavior in the
    thermodynamic limit. It will always show some \emph{finte size effects}.
    These finite size effects cause a "smearing out" of the phase
    transition. This is stronger for smaller system sizes, as is visible
    in fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM}\footnote{See the appendix \ref{appendix:finiteSizeEffects} for a similar figure for the specific heat.}.
    Clearly the \(L=16\) curve is much less steep than the \(L=128\) curve.\\
    Despite of this one can obtain \(T_c\) by finite size scaling
    methods \cite[S. ??]{NewmanBarkema1999} which also yield the critical
    exponents, which is performed in the next section. Also in \ref{ssec:binderIntersections}
    an easier approach, which yields better values for \(T_{c}\), is presented.

\subsection{Critical Exponents}
    Studies on random lattices with\cite{Lima2000} and without\cite{Janke1994}
    varying coupling constants \(J\) suggest, that the
    critical exponents should not be influenced by the disorder parameter
    \(\sigma\). So they will be used for consistency cross checking and
    comparison with the known exact values \cite[S. 59]{Pelissetto2002}.\\
    % Ausfuehrlichere Herleitung der Skalentheorie, Skalenanalyse
    For a more in depth look at finite size scaling, \cite{Norrenbrock2011}
    offers an explanation.
    It is known, that for large \(L\) near the critical temperature the
    equations \eqref{eq:fsscaling:m}, \eqref{eq:fsscaling:chi} and
    \eqref{eq:fsscaling:g} apply.
    \begin{align}
        \label{eq:fsscaling:m}
        \avg{m_L} &= L^\frac{\beta}{\nu} \tilde{M}\brac{L^\frac{1}{\nu}\brac{T-T_c}}\\
        \label{eq:fsscaling:chi}
        \chi_L    &= L^\frac{\gamma}{\nu} \tilde{C}\brac{L^\frac{1}{\nu}\brac{T-T_c}}\\
        \label{eq:fsscaling:g}
        g          &= \frac{3}{2}\brac{1-\frac{\avg{m^4}}{3\avg{m^2}^2}} \propto \tilde{G}\brac{L^\frac{1}{\nu}\brac{T-T_c}}
    \end{align}
    Where \(g\) from eq. \eqref{eq:fsscaling:g} is a normalized
    Binder cumulant \cite{Binder1981} and \(\tilde{M}, \tilde{C}\) and \(\tilde{G}\)
    are unknown scaling functions. To find the exponents
    \(\beta, \gamma, \nu\) and the critical temperature \(T_c\), one
    varies them until the measured values of the observables collapse on
    one curve. Like the obersables from fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:binder_s_0}
    collapse in fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:collapse_s_0}.
    %\footnote{Examples for a collapse to determine the other critical exponents are in appendix \ref{appendix:finiteSizeScaling}}.
    Note that \(L=16\) is not used for the collapse, because it is a
    rather small value. Eq. \eqref{eq:fsscaling:m}-\eqref{eq:fsscaling:g}
    are approximations for big \(L\). For small \(L\) one needs some
    \emph{corrections to scaling} terms, which are not considered here.\\
    To accomplish the collapse in an semi-automatic and reproduceable
    way with an error estimate, the program
    \texttt{autoscale.py} \cite{autoscale2009} is used.
    \begin{figure}[htbp]
        \centering
        \subfigure[Binder Cumulant $g$][]
        {
            \label{sfig:gettingCrit:binder_s_0}
            \includegraphics[width=0.47\textwidth]{plots/binder_s_0}
        }
        \subfigure[Finite Size Scaling of the Binder Cumulant $g$][]
        {
            \label{sfig:gettingCrit:collapse_s_0}
            \includegraphics[width=0.47\textwidth]{plots/collapse_s_0}
        }
        \caption[Examples of Determining Critical Temperature and Exponents]
        {
            \subref{sfig:gettingCrit:binder_s_0} The Binder cumulant \(g\)
                of an square lattice Ising model (\(\sigma=0\))\\
            \subref{sfig:gettingCrit:collapse_s_0} collapsed by finite
                size scaling
        }
        \label{fig:gettingCrit}
    \end{figure}\\

@@

            exact (\cite[p. 59]{Pelissetto2002}) & \multicolumn{1}{c}{\(0\)} & \multicolumn{1}{c}{\(1\)} & \multicolumn{1}{c}{\(-\frac{7}{4}\)} & \multicolumn{1}{c}{\(\frac{1}{8}\)}\\

@@

            GG           & 0.0 & 0.998(8) & -1.735(2) & 0.1262(4)\\
                         & 0.1 & 0.999(19)& -1.744(5) & 0.133(6) \\
                         & 0.3 & 1.029(30)& -1.724(16)& 0.129(12)\\
                         & 0.5 & 1.006(5) & -1.750(12)& 0.125(13)\\
                         & 1.0 & 1.038(32)& -1.743(17)& 0.123(16)\\

@@

            RNG          & 0.0 & 0.992(11)& -1.740(2) & 0.130(1) \\
                         & 0.1 & 0.987(12)& -1.746(5) & 0.133(4) \\
                         & 0.2 & 1.010(9) & -1.756(14)& 0.123(10)\\
                         & 0.5 & 1.010(16)& -1.750(16)& 0.143(13)\\
                         & 1.0 & 1.013(6) & -1.758(16)& 0.138(13)\\

>>>>

\subsection{Finite Size Effects}
\label{ssec:finitesize}
    The aim of this thesis is to find the critical temperature \(T_c\)
    of the disordered Ising model in dependence of the disorder parameter
    \(\sigma\). At \(T_c\) the mean absolute magnetization \(\avg{|m|}\) of
    the system will show a steep decline to zero and the susceptibility
    \(\chi = \frac{N}{T}\brac{\avg{m^2}-\avg{m}^2}\)
    will diverge. In fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM_L}
    it is easy to see that the steep decline of \(\avg{|m|}\)
    occurs at lower temperatures \(T\) for higher
    disorder parameters \(\sigma\) (here with a RNG).
    \begin{figure}[htbp]
        \centering
        \subfigure[Dependence of the Phase Transition on $\sigma$][]
        {
            \label{sfig:smeared_out:meanM_L}
            \includegraphics[width=0.45\textwidth]{plots/Mean_M_L_128}
        }
        \subfigure[Example for Finite Size Effects][]
        {
            \label{sfig:smeared_out:meanM}
            \includegraphics[width=0.45\textwidth]{plots/meanM}
        }
        \caption[Phase Transition and Finite Size Effects]
        {
            \subref{sfig:smeared_out:meanM_L}: Effects of the disorder
            parameter $\sigma$ on the phase transition
            for an underlying RNG with $L=128$.\\
            \subref{sfig:smeared_out:meanM}: Effects of different system
            sizes at \(\sigma = 0\). Dotted lines are guides to the eye.
        }
        \label{fig:smeared_out}
    \end{figure}\\
    But in the shown plots there occurs no steep decline to zero, but a
    smooth decline. \(\avg{\abs{m}}(T_c) = 0\) is only present in infinite
    systems, hence no computer simulation will show the exact behavior in the
    thermodynamic limit. It will always show some \emph{finte size effects}.
    These finite size effects cause a "smearing out" of the phase
    transition. This is stronger for smaller system sizes, as is visible
    in fig. \ref{fig:smeared_out}\subref{sfig:smeared_out:meanM}\footnote{See the appendix \ref{appendix:finiteSizeEffects} for a similar figure for the specific heat.}.
    Clearly the \(L=16\) curve is much less steep than the \(L=128\) curve.\\
    Despite of this one can obtain \(T_c\) by finite size scaling
    methods \cite[S. ??]{NewmanBarkema1999}, which also yield the critical
    exponents. This will be performed in the next section.
    Though there is an easier approach, which yields even more
    precise values for \(T_{c}\), which is presented in \ref{ssec:binderIntersections}.

\subsection{Critical Exponents}
    The critical exponents define the course of an observable near its
    divergence. For instance the vicinity of the divergence of the susceptibility
    can be approximated by
        \chi \propto \abs{T-T_{c}}^{-\beta}
        \label{eq:critExp:beta}
    As the name critical \emph{exponent} suggests, this will often
    be a power law. But not obligatory e.g.\ the critical "exponent" \(\alpha\)
    corresponding to the divergence of the specific heat \(c\) for the two dimensional
    Ising model diverges logarithmic and is nominally \(\alpha = 0\).
    Because of this, \(\alpha\) will not be considered in this section.\\
    The important property of the critical exponents is, that they are
    universal for a model i.e.\ they are independent of the size, the structure
    or the coupling constant \(J\) \cite{NewmanBarkema1999} of the lattice,
    therefore they should also be independent of \(\sigma\).
    Studies on random lattices with\cite{Lima2000} and without\cite{Janke1994}
    varying coupling constants \(J\) confirm, that the
    critical exponents are not influenced by an random lattice.
    So they will be used for consistency cross checking and
    comparison with the known exact values \cite[S. 59]{Pelissetto2002}.
    If the critical exponents can be reproduced, the determined critical
    temperatures are probably correct too.\\
    % Ausfuehrlichere Herleitung der Skalentheorie, Skalenanalyse
    To determine the critical exponents, the finite size scaling method
    will be used. The here given explanation will give a rough idea, how
    the method works, for a more in depth explanation look at \cite{Norrenbrock2011} or \cite{NewmanBarkema1999}.
    The crucial element is that there exists a \emph{scaling function},
    which is valid for sufficiently large \(L\) near the critical
    temperature \(T_{c}\) and has only an explicit \(L\) dependence.
    One can then express some observables like in eq. \eqref{eq:fsscaling:m}, \eqref{eq:fsscaling:chi} and
    \eqref{eq:fsscaling:g}.
    \begin{align}
        \label{eq:fsscaling:m}
        \avg{m_L} &= L^{-\frac{\beta}{\nu}} \tilde{M}\brac{L^\frac{1}{\nu}\brac{T-T_c}}\\
        \label{eq:fsscaling:chi}
        \chi_L    &= L^{\frac{\gamma}{\nu}} \tilde{C}\brac{L^\frac{1}{\nu}\brac{T-T_c}}\\
        \label{eq:fsscaling:g}
        g          &= \frac{3}{2}\brac{1-\frac{\avg{m^4}}{3\avg{m^2}^2}} \propto \tilde{G}\brac{L^\frac{1}{\nu}\brac{T-T_c}}
    \end{align}
    Where \(g\) from eq. \eqref{eq:fsscaling:g} is a normalized
    Binder cumulant \cite{Binder1981} and \(\tilde{M}, \tilde{C}\) and \(\tilde{G}\)
    are the scaling functions.
    To find the exponent e.g. \(\nu\), one takes \eqref{eq:fsscaling:g},
    solves for \(\tilde{G}\), adjusts the axis to represent \(y=\tilde{G}(x)\),
    and plots the measured observables for all \(L\).
    Then one varies \(\nu\) and \(T_{c}\) until the plotted observables
    collapse on one curve -- the scaling function.
    Like the obersables from fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:binder_s_0}
    collapse in fig. \ref{fig:gettingCrit}\subref{sfig:gettingCrit:collapse_s_0}.
    %\footnote{Examples for a collapse to determine the other critical exponents are in appendix \ref{appendix:finiteSizeScaling}}.
    The same principle can be used to determine the other two exponents.
    Note that \(L=16\) is not used for the collapse, because it is a
    rather small value. Eq. \eqref{eq:fsscaling:m}-\eqref{eq:fsscaling:g}
    are approximations for big \(L\). For small \(L\) one needs some
    \emph{corrections to scaling} terms, which are not considered here.\\
    To accomplish the collapse in an semi-automatic and reproduceable
    way with an error estimate, the program
    \texttt{autoscale.py} \cite{autoscale2009} is used.
        \subfigure[Binder Cumulant $g$][]
            \label{sfig:gettingCrit:binder_s_0}
            \includegraphics[width=0.47\textwidth]{plots/binder_s_0}
        \subfigure[Finite Size Scaling of the Binder Cumulant $g$][]
            \label{sfig:gettingCrit:collapse_s_0}
            \includegraphics[width=0.47\textwidth]{plots/collapse_s_0}
            \subref{sfig:gettingCrit:binder_s_0} The Binder cumulant \(g\)
            \subref{sfig:gettingCrit:collapse_s_0} collapsed by finite
        \label{fig:gettingCrit}

@@

            exact (\cite[p. 59]{Pelissetto2002}) & \multicolumn{1}{c}{\(0\)} & \multicolumn{1}{c}{\(1\)} & \multicolumn{1}{c}{\(\frac{7}{4}\)} & \multicolumn{1}{c}{\(\frac{1}{8}\)}\\

@@

            GG           & 0.0 & 0.998(8) & 1.735(2) & 0.1262(4)\\
                         & 0.1 & 0.999(19)& 1.744(5) & 0.133(6) \\
                         & 0.3 & 1.029(30)& 1.724(16)& 0.129(12)\\
                         & 0.5 & 1.006(5) & 1.750(12)& 0.125(13)\\
                         & 1.0 & 1.038(32)& 1.743(17)& 0.123(16)\\

@@

            RNG          & 0.0 & 0.992(11)& 1.740(2) & 0.130(1) \\
                         & 0.1 & 0.987(12)& 1.746(5) & 0.133(4) \\
                         & 0.2 & 1.010(9) & 1.756(14)& 0.123(10)\\
                         & 0.5 & 1.010(16)& 1.750(16)& 0.143(13)\\
                         & 1.0 & 1.013(6) & 1.758(16)& 0.138(13)\\
